Bounded Model Checking for Embedded Systems \\ (Position Paper)

### 2016-01-04 paper.tex
minor changes

State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages. In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness. For instance, such aspects are of extreme relevance in micro-grids and cyber-physical system, in order to ensure reliability, which is a key issue for (smart) cities, industries and consumers.
%In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification. These aspects are of extreme relevance in micro-grids to ensure reliability, which is a key issue for (smart) cities and consumers.

@@

In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the theories supported by SMT solvers ({\it e.g.}, real-time, memory allocation, interrupts, and concurrency). Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}. Consequently, they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth. In particular, BMC techniques limit the addressed regions of data structures ({\it e.g.}, arrays) and the number of loop iterations to a given bound $k$, which then restricts the state space that must be explored during verification, in such a way that real errors in applications \cite{Clarke04,MerzFS12,Ivancic05,Cordeiro12} can be found. Nonetheless, BMC tools are susceptible to exhaustion of time or memory limits, when checking programs with loops whose bounds are too large or cannot be statically determined.

@@

One technique typically used to prove properties, for any given depth, is mathematical induction. In particular, the algorithm called \textit{k}-induction was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints specified during the system design phase~\cite{EenS03}. Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}. Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a new algorithm for proving correctness of (a large set of) C programs by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}. However, there is still little evidence in the associated literature that model checking embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.

@@

It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically addresses programing language issues and general correctness, without taking into account target applications or system behavior. Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if source code is correctly written, but also if it will respond properly in real environments or under external problems. For instance, the anti-lock braking system software of a car model can be bug free, but it may not work properly if a sensor is damaged.

>>>>

State-of-the-art verification methodologies for embedded systems generate test vectors (with constraints), use assertion-based verification, and high-level processor models during simulation~\cite{Behrend15,Lettnin09}; however, embedded system verification also raises additional challenges, such as time and energy constraints, handling of concurrent software, platform restrictions, and legacy designs, which are usually developed in low-level languages. In summary, the main challenges regarding the verification of embedded systems lie on improving coverage, where more system's functions are verified, reducing verification time, {\it i.e.}, pruning the state-space exploration during verification, and incorporating system models, which allows specific checks regarding system behavior and not only code correctness. For instance, such aspects are of extreme relevance in micro-grids and cyber-physical systems, in order to ensure reliability, which is a key issue for (smart) cities, industries, and consumers.
%In summary, the main challenge when verifying embedded systems lies on improving coverage, where more system's functions are verified, and reducing verification time, {\it i.e.}, pruning the state-space exploration during verification. These aspects are of extreme relevance in micro-grids to ensure reliability, which is a key issue for (smart) cities and consumers.

@@

In order to reason about embedded software accurately, an SMT-based BMC must consider a number of issues that are not easily mapped into the (background) theories supported by SMT solvers, {\it e.g.}, real-time, memory allocation, interrupts, and concurrency. Moreover, BMC techniques are able to falsify properties up to a depth \textit{k}. Consequently, they are able to prove system correctness only if an upper bound \textit{k} is known, {\it i.e.}, a bound that unfolds all loops and recursive functions to their maximum possible depth. In particular, BMC techniques limit the addressed regions of data structures ({\it e.g.}, arrays) and the number of loop iterations to a given bound $k$, which then restricts the state space that must be explored during verification, in such a way that real errors in applications \cite{Clarke04,MerzFS12,Ivancic05,Cordeiro12} can be found. Nonetheless, BMC tools are susceptible to exhaustion of time or memory limits, when checking programs with loops whose bounds are too large or cannot be statically determined.

@@

A feasible alternative to check properties in BMC is to prove that an invariant (assertion) is \textit{k}-inductive~\cite{EenS03,Sheera00}. 
The simplest form of \textit{k}-induction consists of two steps: the \textit{base-case} and the \textit{induction-step}. In the base-case, we aim to check that $\phi$ holds in all states reachable from an initial state within $k$ steps (we assume that $k \geq 0$) and in the induction-step, we aim to check that whenever $\phi$ holds in $k$ consecutive states $s_1,\ldots,s_k$, $\phi$ also holds in the next state $s_{k+1}$ of the system. In particular, the \textit{k}-induction algorithm was successfully applied to ensure that (restricted) C programs do not contain data races~\cite{Donaldson10,Kinductor} and do respect time constraints specified during the system design phase~\cite{EenS03}. Additionally, the \textit{k}-induction is a well-established technique in hardware verification, where it is easier to be applied, due to the monolithic transition relation present in such designs~\cite{EenS03,GrosseLD09,Sheera00}. Recently, Gadelha {\it et al.} and Beyer {\it et al.} presented a new algorithm for proving correctness of (a large set of) C programs by mathematical induction, in a completely automatic way ({\it i.e.}, users do not need to provide the loop invariant)~\cite{Gadelha15,Beyer15}. However, there is still little evidence in the associated literature that model checking embedded software, using \textit{k}-induction (and invariants), can be efficiently exploited in embedded system verification~\cite{Rocha15,Brain15}.

@@

It is worth noticing that, currently, SMT-based BMC approaches check for code properties in real programs, which basically address programing language issues and general correctness, without taking into account target applications or system behavior. Such a statement is important, since, as already mentioned, many system features are being moved to software domain, which then claims for schemes that not only check if source code is correctly written, but also if it will respond properly in real environments or under external problems. For instance, the anti-lock braking system software of a car model can be bug free, but it may not work properly if a sensor is damaged.
